FIXME

### 2013-09-25 writeups/proposal/proposal.tex
Wed commits
Former-commit-id: 8d0a7b06f11f1ef3ca5f2f6b6033b416509c1333 [formerly 44f691c9e6f2e8a528aed1ceccf3a197dfb55ee3]
Former-commit-id: 433372b72146985b9a273f575872e5f401eb5d5c

A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
This is similar to singular value decomposition in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance. 
Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
The LDA model performed well, and generated relevant topics.

However, a topic-generating model is not performing the same task as a hashtag-prediction model.
The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
It is unclear though if a user would want to label their tweet with these general topics.
Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
Now it is certainly the case that the topics generated by a latent-based model may be similar, if not identical, to the hashtags used in the post.
However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
Why not take those topics already present in the data and build a model that produces them?
In other words, the topics data are already in the tweet, so why not use them?

@@

Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
I am curious to see the prediction difference between a vector-based models and current co-occurrence models on the StackOverflow dataset.
Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, I am curious to see how a cognitively-plausible vector-based model performs at the task.

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that the few words used just before a hashtag embedded in a tweet will be highly predictive of the hashtag chosen.
To our knowledge, only bag-of-words models have been tested on Twitter to predict hashtags, where co-occurrence between words and hashtags are computed independently from where the word is located in the tweet.
It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.

@@

ACT-R \parencite{Anderson2004} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
The equations that make up this system can be described through a rational analysis of long-term memory retrieval.
That is, given the task of retrieving a chunk of information from long-term declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior number of times it has been needed.
Those prior probabilities for each chunk are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and current environment state.

@@

A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACT-RModel}.
The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability that a chunk will be retrieved from memory increases with total activation ($A_{i}$).

@@

For some domains, it is reasonable to assume that the base-level activations of each chunk within a particular window of interest are stable, and do not change within that window.
Programming-language popularity over the past few years is a reasonable example of this.
Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., \emph{closure}'s growth), it is not the case that the changes have been drastic.
\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years, etc.

Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of being exposed to a chunk.
For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.

@@

Using this assumption for large datasets, the interpretation of the context-adjusted odds ratio becomes much simpler.
If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
If this ratio is less than one, then the log is negative, which provides a negative adjustment of total activation since this context has been observed more often in general than with this particular tag.

@@

Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.

@@

The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford WebBase Project \parencite{Farahat2004}.
So this implementation did not fit into RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.

The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
The links that user chose in experimental testing were largely predictive of the strength of association measure.
The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.

@@

A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
The activation of each tag ($A_{i}$) is a function of four components:
That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy metric was used to compute the attentional weight ($W_{j}$).

The offset term was added to normalize the mean activation across retrievals.
This allowed the use of a logistical regression statistical technique to calibrate the weights for each of the terms.
The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
The entropy metric was used to compute the attentional weight for two reasons:
Stop words could be included in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
Using an entropy metric (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
The process starts with a set of N documents.
A word x document co-occurrence matrix is built by counting the number of times each word appears in each document.
The dimensionality of the full word frequency co-occurrence matrix is then reduced and forced to be represented on a smaller set of dimensions.
This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space, which would be similar to computing the cosine of two correlated eigenvectors.

Usually information is lost when the data are restructured into a smaller set of latent dimensions.
However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
They tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
Model performance also improved when the original dimensionality of the data was reduced.

@@

When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.

@@

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
For example, assume that a user has written the word ``zend'' in a post.
``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP}.
Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
So the cosine between the environment vector \emph{PHP} and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.

@@

Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is an example where the encoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

@@

For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
Within this corpus, the phrase ``martin luther'' often occurred.
When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.

@@

A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $[1,-1]$s in random locations. 
In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is bit shifted to the left one position ($^{-1}$).
This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.

@@

Random permutations use one of simplest permutations possible to create a new environment vector.
This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
However, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.

@@

For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular column in the vector.
If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
However, each word's environment vector is compressed by using a length much less than the number of unique words and using a combination of a few [1,-1]s to represent the word.

@@

Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
It may be interesting to explore how model performance changes when one calculation is swapped out for another.
For example, a vector-based model could be used to generate a compress (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
It's unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.

>>>>

A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance. 
Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
The LDA model performed well, and generated relevant topics.
When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).

However, a topic-generating model is not performing the same task as a hashtag-prediction model.
The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
It is unclear though if a user would want to label their tweet with these general topics.
Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
Now it is certainly the case that the topics generated by a latent-based model may be similar to the hashtags used in the post.
However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
Why not take those topics already present in the data and build a model that produces them?
In other words, the topics data are already in the tweet, so why not use them?

@@

Although recent work has been done to model human tagging on the StackOverflow site, to our knowledge no vector-based holographic memory model has been tested or used.
I am curious to see the prediction difference between a vector-based model and current co-occurrence models on the StackOverflow dataset.
Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.

@@

ACT-R \parencite{Anderson2004} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
The equations that make up this system can be derived through a rational analysis of long-term memory retrieval.
That is, given the task of retrieving a chunk of information from long-term declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of being retrieved again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adusted by current environment state.

@@

A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACT-RModel}.
The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability ($P_{i}$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).

@@

For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
Programming-language popularity over the past few years is a reasonable example of this.
Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., \emph{closure}'s growth), it is not the case that the changes have been drastic.
\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years.

Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of being exposed to that chunk compared to all other chunk exposures.
For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.

@@

Using this assumption for large datasets ($S_{ji} = log \frac{p(i|j)}{p(i)}$), the interpretation of the context-adjusted odds ratio becomes much simpler.
If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
If this ratio is less than one, then the log is negative, which provides a negative adjustment of total activation since this context has been observed more often in general than with this particular tag.

@@

Further, this PMI equation is actually identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.
That is, the PMI equation \emph{is} the simplified form of ACT-R's strength of association index ($S_{ji}$).

@@

The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford WebBase Project \parencite{Farahat2004}.
So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.

%TODO: This paragraph doesn't fit well after the previous one, and in the scaling section
%The SNIF-ACT model uses strength of association to measure the relationship between the words in a search query and the words contained in each query result.
%A higher $S_{ji}$ between a query and a result means that the result is a better fit for the query, and that the user is more likely to click on that link.
%The links that user chose in experimental testing were largely predictive of the strength of association measure.
%The predictive power of the SNIF-ACT model went above and beyond what would be predicted by simply assuming that the user always clicks on the first link in the search results page.

@@

A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
The activation of each tag ($A_{i}$) is a function of four components:
That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).

The offset term was added to normalize the mean activation across retrievals.
This allowed the use of a logistical regression statistical technique to calibrate the weights for each of the terms.
The title and body contextual components were separated since it seemed plausible that the words in the title would be better predictors of the tag used than the words in the body.
It actually turned out that the words in the body were a better predictor of tag use \parencite{Stanley2013}, most likely because there are many more words in the body than the title of the post.
The entropy measure was used to compute the attentional weight for two reasons:
Stop words could be statistically derived and attentuated in the analysis, and more predictive words of some tag (e.g., the word ``PHP'' in a post) were given higher weight.
Using an entropy measure (as opposed to weighting each word equally in the post) allows the model to allocate limited attentional resources ($W$) to important contextual elements.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
The process starts with a set of N documents.
A word x document co-occurrence matrix is built by counting the number of times each word appears in each document.
The dimensionality of the full word frequency co-occurrence matrix is then reduced and forced to be represented on a smaller set of dimensions.
This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

Usually information is lost when the data are restructured into a smaller set of latent dimensions.
However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
They tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
Afterwards further reductions in the dimensionality of the data began to degrade performance.

@@

When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.

@@

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
For example, assume that a user has written the word ``zend'' in a post.
``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{PHP}$).
Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
So the cosine between the environment vector \emph{PHP} ($e_{PHP}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.

@@

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is an example where both the encoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

@@

For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
Within this corpus, the phrase ``martin luther'' often occurred.
When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-forward location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-back location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.

@@

A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few $<1,-1>$s in random locations. 
In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector ($e_{i}$) is bit shifted to the left one position ($^{-1}$).
This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.

@@

Random permutations usees one of simplest permutations possible to create a new environment vector.
This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
However, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.

@@

For random permutations, each tag is represented by a compressed vector with a length much less than the number of words.
Note that if the length of a word's environment vector was the number of unique words, then each word could be represented by associating it with a particular row in the column vector.
If done in this way, building a matrix of all words' environment vectors x tags would recover the original word x document matrix \parencite{Kanerva2000}.
However, each word's environment vector is compressed using a length much less than the number of unique words and using a combination of a few $<1,-1>$s to represent the word.

@@

Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
It may be interesting to explore how model performance changes when one calculation is swapped out for another.
For example, a vector-based model could be used to generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
It's unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.
